### 2012-08-03 Doc/VRIPHYS2012/vriphys2012.tex
adding paper outline to introduction
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@453 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize -- Submission ID
%\author[paper1026]
%       {paper1026}
\author[T. Golembiovský \& C. Duriez]
       {Tomáš Golembiovský$^{1,2}$ and Christian Duriez$^{2}$\\
        $^1$Faculty of Informatics, Masaryk University, Czech Republic\\
        $^2$INRIA Lille, France
       } % TODO: ask Christian how to specify affiliation

@@

% TODO: paper structure
TODO: paper structure

@@

Because the deflection field for the element is non-linear \eqref{eq-bez-delta}
in position the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 6-point Gaussian
quadrature for integration over triangle area.
TODO: figure + gives the formula of the quadrature...
% TODO: add ref
% TODO: add abscissas/weights?

@@

Which means we use the same expression to interpolate the velocity only by
substituting the velocities at control points. 
For the corner control points  we
already have the velocities $V_i$. For the internal control points  we need to compute
them. We again make use of the kinematic link between control points and control points.
%solve the problem of computing the velocity of a point attached to the moving rigid body.
For the edge points we applied to motion of a rigid body:
%
\begin{eqnarray}
    V_j = V_i + \omega_i \times P_{i-j}
\end{eqnarray}
%
where $i$ is the index of the corner node the edge node is attached to and
$\omega_i$ is the angular velocity at the node. For the central node
the formula is based on the derivate of equation \eqref{eq-central2}:
%
\begin{eqnarray}
    V_{10} = \frac{1}{3} \sum_{i=1}^3 \left( V_i + \omega_i \times P_{i-10} \right)
\end{eqnarray}

@@

In figure \ref{fig-fps} we present performance of the element in terms of
frames per second. All tests were performed on machine equipped with Dual-Core
AMD Opteron Processor 2218 and 3~GB of RAM. We used a conjugate gradient solver
to solve the system. Our implementation is without optimizations
and isn't parallel, which means only single CPU core was used for the
simulation. The values reported are for the raw physical simulation with (b)
and without (a) the frame fixing method described in subsection
\ref{sec-corot}. We also present results including the mapping of
high-resolution mesh onto the mechanical mesh for visualisation purposes or
collisions.

@@

The planning process can be improved by employing surgical simulation system
using three-dimensional heart models based on patient-specific images as input.
Most present simulation systems focus on execution of low-level surgical
procedures, e.g. incising and suturing. In these reactive simulation systems it
is more important to enable simulations in real-time than accurate simulation
of tissue and blood vessel deformation in response to simulated interventions.

To be able to predict results of complex surgical procedures a physically more
accurate approach is necessary. Using our shell finite element model we have
developed a prototype \cite{Kislinskiy2012} of new surgical simulation system
for preoperative planning. The element has been used to model deformations of
blood vessels and artificial patches. By performing a set of topological changes
we simulate the process of incising the vessel and suturing on a pre-specified
boundary. By iteratively adjusting the underlaying mesh we search for the
elastic equilibrium. Figure \ref{fig-chd} presents a screenshots of the
application prototype.

@@

Currently, due to the technical decisions the element has two limitations.
The first limitation is that the element does not provide $C^1$ continuity
across edges. It is $C^1$ continuous at corners but only $C^0$ on the
edges. This may prove to be an issue for example if curvature needs to be
computed not only inside the element but also across the edges.

@@

Because of the formulation of B\'ezier triangle is intrinsically present
there is a clean and obvious way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation.
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.

>>>>

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% TODO: anonymize -- Submission ID
       \author[paper1026]
              {paper1026}
%\author[T. Golembiovský \& C. Duriez]
%       {Tomáš Golembiovský$^{1,2}$ and Christian Duriez$^{2}$\\
%        $^1$Faculty of Informatics, Masaryk University, Czech Republic\\
%        $^2$INRIA Lille, France
%       } % TODO: ask Christian how to specify affiliation

@@

We start by reviewing the definition of B\'ezir triangle. Then the construction
of initial control mesh is described and  kinematic link between control points
is defined. In Section 3 the FE model is derived and the method of computing
the corotational frame is described. In Section 4 we specify how to map a high
resolution surface mesh onto the mechanical mesh and give formulas for
computing velocities of points on the surface and projecting of forces. Section
5 presents test of convergence, speed analysis and two created applications
that use the described shell model. We conclude with discussion of present
limitations.

@@

Because the deflection field for the element is non-linear \eqref{eq-bez-delta}
in position the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 6-point Gaussian
quadrature for integration over triangle area.
%TODO: figure + gives the formula of the quadrature...
% TODO: add ref
% TODO: add abscissas/weights?

@@

For the corner control points  we
already have the velocities $V_i$. For the internal control points  we need to compute
them. We again make use of the kinematic link between control points and control points.
%solve the problem of computing the velocity of a point attached to the moving rigid body.
For the edge points we applied to motion of a rigid body:
%
\begin{eqnarray}
    V_j = V_i + \omega_i \times P_{i-j}
\end{eqnarray}
%
where $i$ is the index of the corner node the edge node is attached to and
$\omega_i$ is the angular velocity at the node. For the central node
the formula is based on the derivate of equation \eqref{eq-central2}:
%
\begin{eqnarray}
    V_{10} = \frac{1}{3} \sum_{i=1}^3 \left( V_i + \omega_i \times P_{i-10} \right)
\end{eqnarray}

@@

In figure \ref{fig-fps} we present performance of the element in terms of
frames per second. All tests were performed on machine equipped with Dual-Core
AMD Opteron Processor 2218 and 3~GB of RAM. We used a conjugate gradient solver
to solve the system. Our implementation is without optimizations
and only single CPU core was used for the
simulation. The values reported are for the raw physical simulation with (b)
and without (a) the frame fixing method described in subsection
\ref{sec-corot}. We also present results including the mapping of
high-resolution mesh onto the mechanical mesh for visualisation purposes or
collisions.

@@

To be able to predict results of complex surgical procedures a surgical simulation system can be employed.
We have
developed a prototype \cite{Anonymous2012} of new surgical simulation system
for preoperative planning. The element has been used to model deformations of
blood vessels and artificial patches. By performing a set of topological changes
we simulate the process of incising the vessel and suturing on a pre-specified
boundary. By iteratively adjusting the underlaying mesh we search for the
elastic equilibrium.

@@

Currently, due to the technical decisions the element has two limitations.
It is $C^1$ continuous at corners but only $C^0$ on the
edges. This may prove to be an issue for example if curvature needs to be
computed not only inside the element but also across the edges.

@@

The formulation of B\'ezier triangle provides
a clean way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation.
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.
